---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# boiwsa

<!-- badges: start -->
<!-- badges: end -->

`boiwsa` is an R package for performing weekly seasonal adjustment on time series data. It provides a simple, easy-to-use interface for calculating seasonally adjusted estimates of weekly data, as well as a number of diagnostic tools for evaluating the quality of the adjustments.

The seasonal adjustment procedure is based on a locally-weighted least squares procedure (Cleveland et al., 2014).


We consider the following decomposition model:

$$
y_{t}=T_{t}+S_{t}+H_{t}+O_{t}+I_{t},
$$

where $T_{t}$, $S_{t}$ , $O_{t},$ $H_{t}$ and $I_{t}$ represent the trend,
seasonal, outlier, holiday- and trading-day, and irregular components,
respectively. The seasonal component is modeled as


$$\begin{eqnarray*}
S_{t} &=&\sum_{k=1}^{K}\left( \alpha _{k}^{y}\sin (\frac{2\pi kD_{t}^{y}}{
n_{t}^{y}})+\beta _{k}^{y}\cos (\frac{2\pi kD_{t}^{y}}{n_{t}^{y}})\right) +
\\
&&\sum_{l=1}^{L}\left( \alpha _{l}^{m}\sin (\frac{2\pi kD_{t}^{m}}{n_{t}^{m}}
)+\beta _{l}^{m}\cos (\frac{2\pi kD_{t}^{m}}{n_{t}^{m}})\right) ,
\end{eqnarray*}$$


where $D_{t}^{y}$ and $D_{t}^{m}$ are the day of the year and the day of the
month, and $n_{t}^{y}$ and $n_{t}^{m}$ are the number of days in the given
month or year. Thus, the seasonal adjustment procedure takes into account the existence of two cycles, namely intrayearly and intramonthly.



The trend component is extracted with Friedman's SuperSmoother using `stats::supsmu()`. 





## Installation

To install boiwsa, you can use devtools:

``` r
# install.packages("devtools")
devtools::install_github("timginker/boiwsa")
```

Alternatively, you can clone the repository and install the package from source:

``` bash
git clone https://github.com/timginker/boiwsa.git
cd boiwsa
R CMD INSTALL .
```



## Usage

Using `boiwsa` is simple. First, load the `boiwsa` package:

```{r}
library(boiwsa)
```

Next, load your time series data into a data frame object. Here is an example that is based on the `gasoline` data from the `fpp2` package:


```{r}
data("gasoline.data")

plot(gasoline.data$date,gasoline.data$y,type="l",xlab="Year",ylab=" ", main="Weekly US gasoline production")
```


Once you have your data loaded, you can use the `boiwsa` function to perform weekly seasonal adjustment:

```{r}

res=boiwsa(x=gasoline.data$y,dates=gasoline.data$date)

```

The `x` argument takes the series to be seasonally adjusted, while the dates argument takes the associated dates in date format. Unless specified otherwise (i.e., `my.k_l = NULL`), the procedure automatically identifies the best number of trigonometric variables to capture the yearly ($K$) and monthly ($L$) cycles based on the AICc. The information criterion is specified by the `ic` option.

The procedure automatically searches for additive outliers (AO) using the method described in Appendix C of Findley et al. (1998). To disable the automatic AO search, set `auto.ao.search = F`. To add user-defined AOs, use the `ao.list` option.


The `boiwsa` function returns a list object containing the results. The seasonally adjusted series is stored in a vector called `sa`. In addition, the estimated seasonal factors are stored as `sf`. 


You can then plot the adjusted data to visualize the seasonal pattern:

```{r}
plot(gasoline.data$date,gasoline.data$y,type="l",xlab="Year",ylab=" ", main="Weekly US gasoline production")
lines(gasoline.data$date,res$sa,col="red")
legend(
  "topleft",
  legend = c("Original", "SA"),
  lwd = c(2,2),
  col = c("black", "red"),
  bty = "n"
)
```


To evaluate the quality of the adjustment, you can use the `plot_spec` function provided by the package, which generates a plot of the autoregressive spectrum of the raw and seasonally adjusted data:



```{r}
plot_spec(res)
```



# References

Cleveland, W.P., Evans, T.D. and S. Scott (2014). Weekly Seasonal Adjustment-A Locally-weighted Regression Approach (No. 473). Bureau of Labor Statistics.



Findley, D.F., Monsell, B.C., Bell, W.R., Otto, M.C. and B.C Chen (1998). New capabilities and methods of the X-12-ARIMA seasonal-adjustment program. Journal of Business & Economic Statistics, 16(2), pp.127-152.


# Disclaimer

The views expressed here are solely of the author and do not necessarily represent the views of the Bank of Israel. 

Please note that `boiwsa` is still under development and may contain bugs or other issues that have not yet been resolved. While we have made every effort to ensure that the package is functional and reliable, we cannot guarantee its performance in all situations. 

We strongly advise that you regularly check for updates and install any new versions that become available, as these may contain important bug fixes and other improvements. By using this package, you acknowledge and accept that it is provided on an "as is" basis, and that we make no warranties or representations regarding its suitability for your specific needs or purposes.


